\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage[
  backend=biber,
  style=numeric-comp,
  sorting=none
]{biblatex}
\addbibresource{refs.bib}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\CZ}{\mathcal{C}_{\mathrm{ZAP}}}
\newcommand{\Zb}{Z_{\mathrm{base}}}
\newcommand{\Zh}{Z_{\mathrm{http}}}
\newcommand{\Zw}{Z_{\mathrm{ws}}}
\newcommand{\Zf}{Z_{\mathrm{fix}}}
\newcommand{\Zr}{Z_{\mathrm{rns}}}
\newcommand{\Zm}{Z_{\mathrm{mcp}}}
\newcommand{\Zac}{Z_{\mathrm{acp}}}
\newcommand{\Za}{Z_{\mathrm{a2a}}}
\newcommand{\Adv}{\mathbf{Adv}}

\title{Compositionality of the ZAP Protocol Family:\\Formal Proofs}
\author{ZAP Protocol Authors\\\texttt{zap-proto.io}}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
This document gives the formal proofs underlying the categorical
framework presented in the companion paper. We prove the category
axioms for $\CZ$, the initiality of $\Zb$, the security composition
theorem under tight embeddings, and the tightness of the five
concrete embeddings claimed in the family.
\end{abstract}

\tableofcontents

\section{Preliminaries}
\label{sec:prelim}

We adopt the protocol model of the companion paper. A protocol $P$
is the tuple $(\Sigma_P, \mathcal{S}_P, s^0_P, \to_P,
\mathcal{V}_P)$. We allow $P$ to be probabilistic by lifting $\to_P$
to a sub-distribution-valued transition function
$\mu_P : \mathcal{S}_P \times \Sigma_P \to \mathcal{D}(\mathcal{S}_P)$
where $\mathcal{D}$ denotes finite sub-distributions. We assume the
underlying probability space is fixed and standard measure-theoretic
care is taken throughout.

We write $|m|$ for the bit length of a wire message $m$ and
$|\sigma| = \sum_i |m_i|$ for the total bit length of a trace
$\sigma$. The security parameter is $\lambda \in \mathbb{N}$, and a
function $\nu : \mathbb{N} \to \mathbb{R}_{\ge 0}$ is
\emph{negligible}, written $\nu \in \mathrm{negl}(\lambda)$, if for
every polynomial $p$ there is $\lambda_0$ such that $\nu(\lambda) <
1/p(\lambda)$ for all $\lambda \ge \lambda_0$.

We use the standard fact that statistical distance
$\Delta(X, Y) := \frac{1}{2}\sum_{z}|\Pr[X = z] - \Pr[Y = z]|$ is a
metric on the space of distributions over a discrete set.

\section{Theorem 1: Category and 2-Category Axioms}
\label{sec:cat-axioms}

\begin{theorem}[Category and 2-category axioms for $\CZ$]
\label{thm:cat-axioms}
The data $(\mathrm{Obj}(\CZ), \mathrm{Hom}_{\CZ},
\mathrm{2Hom}_{\CZ}, \circ, \mathrm{id})$ as defined in the
companion paper forms a strict 2-category. Specifically:
\begin{enumerate}
\item For every object $P$, the identity $\mathrm{id}_P : P
\hookrightarrow P$ is an embedding.
\item Composition of embeddings is associative:
$(\chi \circ \psi) \circ \varphi = \chi \circ (\psi \circ \varphi)$
for all composable $\varphi, \psi, \chi$.
\item Identity is left and right unit for composition:
$\mathrm{id}_Q \circ \varphi = \varphi = \varphi \circ
\mathrm{id}_P$ for $\varphi : P \hookrightarrow Q$.
\item For every parallel pair of embeddings $\varphi, \varphi' : P
\rightrightarrows Q$, the 2-morphisms $\mathrm{2Hom}(\varphi,
\varphi')$ form a category under vertical composition.
\item Horizontal composition of 2-morphisms exists and the
interchange law holds:
$(\beta \cdot \alpha) \circ_h (\beta' \cdot \alpha') =
(\beta \circ_h \beta') \cdot (\alpha \circ_h \alpha')$.
\end{enumerate}
\end{theorem}

\begin{proof}
We treat each axiom in turn.

\paragraph{(1) Identity is an embedding.}
The identity is the triple $(\mathrm{id}_{\Sigma_P}^{(1)},
\mathrm{id}_{\mathcal{S}_P}, 0)$. Each clause of the embedding
definition (Definition~3.1 in the companion paper, restated above) holds
trivially: $\mathrm{id}_{\Sigma_P}^{(1)}$ is injective; the state map
preserves $s^0_P$; transitions are preserved exactly because the
trace map is the singleton; security predicates are preserved
because the state map is the identity; and the leakage profile is
zero by construction.

\paragraph{(2) Associativity.}
Let $\varphi : P \hookrightarrow Q$, $\psi : Q \hookrightarrow R$,
$\chi : R \hookrightarrow S$. We need
$(\chi \circ \psi) \circ \varphi = \chi \circ (\psi \circ \varphi)$
component-wise.

The message components are
$\chi_\Sigma^* \circ (\psi_\Sigma^* \circ \varphi_\Sigma)
= (\chi_\Sigma^* \circ \psi_\Sigma^*) \circ \varphi_\Sigma$, which
follows from associativity of function composition lifted to
sequences (the lifting being natural and preserving composition).

The state components are $\chi_\mathcal{S} \circ (\psi_\mathcal{S}
\circ \varphi_\mathcal{S}) = (\chi_\mathcal{S} \circ
\psi_\mathcal{S}) \circ \varphi_\mathcal{S}$, again by associativity
of function composition.

The leakage components are $\rho_\chi + (\rho_\psi + \rho_\varphi)
= (\rho_\chi + \rho_\psi) + \rho_\varphi$, which follows because
addition over the reals is associative.

\paragraph{(3) Unit laws.}
Both sides reduce, by direct computation, to $\varphi$. The leakage
of identity is zero, $\rho_\varphi + 0 = \rho_\varphi = 0 +
\rho_\varphi$.

\paragraph{(4) Vertical composition of 2-cells.}
Given $\alpha : \varphi \Rightarrow \varphi'$ and $\beta : \varphi'
\Rightarrow \varphi''$ with relations $R_\alpha, R_\beta$, define
the relational composite $R_{\beta \cdot \alpha} := R_\beta \circ
R_\alpha = \{(t, t'') : \exists t', (t, t') \in R_\alpha \wedge
(t', t'') \in R_\beta\}$. We verify the simulation conditions
(Definition~3.4 in the companion paper):

\begin{itemize}
\item For all $s$, $(\varphi_\mathcal{S}(s), \varphi'_\mathcal{S}(s))
\in R_\alpha$ and $(\varphi'_\mathcal{S}(s),
\varphi''_\mathcal{S}(s)) \in R_\beta$, so
$(\varphi_\mathcal{S}(s), \varphi''_\mathcal{S}(s)) \in R_{\beta \cdot
\alpha}$.
\item Whenever $(t, t'') \in R_{\beta \cdot \alpha}$ via witness
$t'$, simulation moves witnessed by $\alpha$ produce a $t_1$ in the
image of $\varphi'$, and simulation moves witnessed by $\beta$
produce a $t''_j$ in the image of $\varphi''$ such that $(t_1,
t''_j) \in R_{\beta \cdot \alpha}$.
\item Observational equivalence composes: the leakage difference
across the composite is $(\rho_{\varphi''} - \rho_{\varphi'}) +
(\rho_{\varphi'} - \rho_\varphi) = \rho_{\varphi''} - \rho_\varphi$
as required.
\end{itemize}

The identity 2-cell on $\varphi$ is the diagonal
$\{(\varphi_\mathcal{S}(s), \varphi_\mathcal{S}(s)) : s \in
\mathcal{S}_P\}$ extended by reachability; identity laws and
associativity follow from the corresponding laws for relational
composition.

\paragraph{(5) Horizontal composition and interchange.}
Given 2-cells $\alpha : \varphi \Rightarrow \varphi'$ in
$\mathrm{Hom}(P, Q)$ and $\beta : \psi \Rightarrow \psi'$ in
$\mathrm{Hom}(Q, R)$, define the horizontal composite
$\beta \circ_h \alpha : \psi \circ \varphi \Rightarrow \psi' \circ
\varphi'$ by the relation
\[
R_{\beta \circ_h \alpha} := \{(\psi_\mathcal{S}(t),
\psi'_\mathcal{S}(t')) : (t, t') \in R_\alpha\}
\cup \{(\psi_\mathcal{S}(u), \psi'_\mathcal{S}(u')) : (u, u') \in
R_\beta \text{ post-action}\}.
\]
The interchange law is a routine diagram chase using the
associativity of relational composition; we omit the calculation as
it parallels the standard 2-category construction
\cite[Ch.~XII.3]{maclane1998categories}.
\end{proof}

\section{Theorem 2: Initiality of $\Zb$}
\label{sec:init-proof}

\begin{theorem}[$\Zb$ is initial in $\CZ$]
\label{thm:init-proof}
For every object $P \in \mathrm{Obj}(\CZ)$, there is exactly one
embedding $\iota_P : \Zb \hookrightarrow P$.
\end{theorem}

\begin{proof}
We construct $\iota_P$ explicitly and then show uniqueness.

\paragraph{Construction.}
Every protocol $P$ in the family is, by specification, layered over
$\Zb$. Concretely, $P$ has an initial state in which it has not yet
processed any $P$-specific envelopes; in this initial phase, the
$\Sigma_P$ messages reduce to bare $\Zb$ envelopes carrying the
handshake. Define:
\begin{itemize}
\item $\iota_{P,\Sigma} : \Sigma_{\Zb} \to \Sigma_P^+$ by
$\iota_{P,\Sigma}(m) := (m)$, treating each $\Zb$ envelope as a
singleton trace of $P$-messages (which the $P$ semantics interprets
as a $\Zb$-only handshake message).
\item $\iota_{P,\mathcal{S}} : \mathcal{S}_{\Zb} \to \mathcal{S}_P$
by $\iota_{P,\mathcal{S}}(s) := (s, \bot)$, where the second
component records that no $P$-specific state has been added.
\item $\rho_{\iota_P} := 0$.
\end{itemize}

We verify the embedding conditions:
\begin{enumerate}
\item Injectivity: $\iota_{P,\Sigma}$ is the identity on the
underlying envelope, so injective.
\item Initial-state preservation: $\iota_{P,\mathcal{S}}(s^0_{\Zb}) =
(s^0_{\Zb}, \bot) = s^0_P$ by construction (the $P$-initial state is
the $\Zb$-initial state with empty $P$-state).
\item Transition preservation: $\Zb$ transitions in the $P$-handshake
phase are by definition $P$-transitions.
\item Security predicate preservation: every property of $\Zb$
present in state $s$ is also present in state $(s, \bot)$ because the
$P$-valuation extends $\mathcal{V}_{\Zb}$ on the $\Zb$-part of the
state.
\item Tightness: $\rho = 0$ trivially satisfies the negligibility
bound.
\end{enumerate}

\paragraph{Uniqueness.}
Suppose $\iota'_P : \Zb \hookrightarrow P$ is another embedding. We
show $\iota'_P = \iota_P$.

The state component must satisfy $\iota'_{P,\mathcal{S}}(s^0_{\Zb}) =
s^0_P = (s^0_{\Zb}, \bot)$. Inductively, if for some $\Zb$ trace
$\sigma$ ending in state $s$ we have
$\iota'_{P,\mathcal{S}}(s) = (s, \bot)$, then for the next $\Zb$
transition $s \xrightarrow{m} s'$, the embedding must produce a
$P$-trace from $(s, \bot)$ to $\iota'_{P,\mathcal{S}}(s')$. Since
$P$ has not yet entered any $P$-specific state, the only available
$P$-transitions from $(s, \bot)$ are those that correspond to $\Zb$
handshake transitions. By transition preservation, this forces
$\iota'_{P,\mathcal{S}}(s') = (s', \bot)$ and the message component
must be the singleton trace.

The leakage component $\rho_{\iota'_P}$ must satisfy
$\rho_{\iota'_P} \le \rho_{\iota_P} = 0$ by minimality (any extra
leakage would arise from extra wire activity, but the message
component is fixed). Hence $\rho_{\iota'_P} = 0$.

We have $\iota'_P = \iota_P$ in all three components. \qedhere
\end{proof}

\begin{remark}[Functoriality of language bindings]
A language binding $L$ defines a category $\mathcal{C}_L$ whose
objects are $L$-typed protocol implementations and whose morphisms
are $L$-typed embeddings. The map $F_L : \CZ \to \mathcal{C}_L$
preserving composition and identity is a functor. By initiality,
the binding type $F_L(\Zb)$ determines the binding types of all
sub-protocols up to isomorphism (each $F_L(P)$ has a unique morphism
from $F_L(\Zb)$ in $\mathcal{C}_L$, namely $F_L(\iota_P)$). This
formalizes wire-level interoperability.
\end{remark}

\section{Theorem 3: Security Composition Under Tight Embedding}
\label{sec:comp-proof}

\begin{theorem}[Security composition under tight embedding]
\label{thm:composition-proof}
Let $\varphi : P \hookrightarrow Q$ be a tight embedding and let
$\pi$ be a security property in $\Pi(\varphi)$ with games
$G_\pi^P, G_\pi^Q$. Suppose $P$ satisfies $\pi$ against adversary
class $\mathcal{C}$ with bound $\epsilon_P(\lambda)$. Then $Q$
satisfies $\pi$ against $\mathcal{C}$ with bound
$\epsilon_Q(\lambda) \le \epsilon_P(\lambda) + \rho_\varphi(\lambda)
+ \delta_\varphi(\lambda)$, where $\delta_\varphi$ is a function of
the embedding's syntactic complexity (specifically, the maximum
length of $\varphi_\Sigma(m)$ over $m \in \Sigma_P$).
\end{theorem}

\begin{proof}
The proof is by reduction. Let $\mathcal{B}$ be a $\mathcal{C}$-class
adversary against $G_\pi^Q$ with advantage $\Adv := \epsilon_Q$. We
construct an adversary $\mathcal{A}$ against $G_\pi^P$ with
advantage at least $\Adv - \rho_\varphi - \delta_\varphi$, which
forces $\Adv \le \epsilon_P + \rho_\varphi + \delta_\varphi$.

\paragraph{Construction of $\mathcal{A}$.}
$\mathcal{A}$ runs $\mathcal{B}$ as a sub-procedure. When
$\mathcal{B}$ asks for the next $Q$-message, $\mathcal{A}$ asks the
$P$-challenger for the corresponding $P$-message and applies
$\varphi_\Sigma$ to obtain a $Q$-trace of length at most $k :=
\max_m |\varphi_\Sigma(m)|$. $\mathcal{A}$ feeds this trace to
$\mathcal{B}$ one message at a time. When $\mathcal{B}$ outputs a
guess $b \in \{0, 1\}$, $\mathcal{A}$ outputs the same guess.

\paragraph{Analysis.}
Let $\mathrm{View}_\mathcal{B}^Q$ be $\mathcal{B}$'s view in the
real $G_\pi^Q$ game and $\mathrm{View}_\mathcal{B}^{Q,\varphi}$ be
its view in the simulated game where the $Q$-trace is
$\varphi_\Sigma$ applied to the $P$-trace. By the tightness
condition, $\Delta(\mathrm{View}_\mathcal{B}^Q,
\mathrm{View}_\mathcal{B}^{Q,\varphi}) \le \rho_\varphi(\lambda)$.

By the standard distinguishing-distance lemma, $\mathcal{B}$'s
advantage in the simulated game is at least $\Adv -
\rho_\varphi(\lambda)$.

The $\delta_\varphi$ term arises from the granularity of the
challenger interaction: $\mathcal{A}$ may need to interleave queries
to the $P$-challenger with sub-message activity by $\mathcal{B}$, and
the resulting timing can introduce a small additional advantage
bounded by $\delta_\varphi(\lambda)$, which depends polynomially on
the maximum embedding length $k$. Specifically, for an embedding
where each $P$-message expands to at most $k$ $Q$-messages and the
adversary interacts with $q$ messages, $\delta_\varphi(\lambda) \le
k \cdot q \cdot 2^{-\lambda}$, which is negligible whenever $k$ and
$q$ are polynomial in $\lambda$.

Therefore $\mathcal{A}$ has advantage at least $\Adv - \rho_\varphi
- \delta_\varphi$ in the $P$-game, but $\mathcal{A} \in \mathcal{C}$
(it does only polynomial work on top of $\mathcal{B}$ and applying
$\varphi_\Sigma$, both of which are polynomial). By assumption,
$\mathcal{A}$'s advantage is at most $\epsilon_P$. Solving for
$\Adv$ gives $\Adv \le \epsilon_P + \rho_\varphi + \delta_\varphi$
as claimed.
\end{proof}

\begin{corollary}[Stack inheritance]
\label{cor:stack-proof}
Let $\varphi_1 : P_1 \hookrightarrow P_2, \ldots, \varphi_n :
P_n \hookrightarrow P_{n+1}$ all be tight, and let $\pi$ be
preserved by every $\varphi_i$. If $P_1$ satisfies $\pi$ with bound
$\epsilon_1$, then $P_{n+1}$ satisfies $\pi$ with bound
\[
\epsilon_{n+1} \le \epsilon_1 + \sum_{i=1}^n
\bigl(\rho_{\varphi_i} + \delta_{\varphi_i}\bigr).
\]
For $n$ polynomial in $\lambda$ and tight embeddings, the bound is
negligible iff $\epsilon_1$ is.
\end{corollary}

\begin{proof}
Induction on $n$.
Base case $n = 0$ is trivial.
Inductive step: assume $P_n$ satisfies $\pi$ with bound
$\epsilon_n \le \epsilon_1 + \sum_{i=1}^{n-1}(\rho_{\varphi_i} +
\delta_{\varphi_i})$. Apply \cref{thm:composition-proof} to
$\varphi_n$ to obtain $\epsilon_{n+1} \le \epsilon_n +
\rho_{\varphi_n} + \delta_{\varphi_n}$, which gives the bound.

For tight embeddings, $\rho_{\varphi_i}, \delta_{\varphi_i} \in
\mathrm{negl}(\lambda)$. A polynomial sum of negligible functions
is negligible, so $\epsilon_{n+1} - \epsilon_1$ is negligible.
\end{proof}

\section{Tightness of Concrete Embeddings}
\label{sec:concrete-proofs}

We prove the five embeddings claimed in the companion paper are
tight.

\subsection{$\varphi_{hw} : \Zh \hookrightarrow \Zw$}

\begin{proposition}[Proposition~5.1 of companion]
\label{prop:hw-proof}
The embedding $\varphi_{hw}$ that maps each $\Zh$ request/response
pair to a $\Zw$ stream opened, used for one round trip, and closed,
is tight.
\end{proposition}

\begin{proof}
We construct the embedding explicitly:
\begin{itemize}
\item $\varphi_{hw,\Sigma}(m_{\mathrm{req}}) = (\mathtt{OPEN},
m_{\mathrm{req}})$ for a $\Zh$ request $m_{\mathrm{req}}$.
\item $\varphi_{hw,\Sigma}(m_{\mathrm{resp}}) = (m_{\mathrm{resp}},
\mathtt{CLOSE})$ for a $\Zh$ response $m_{\mathrm{resp}}$.
\item $\varphi_{hw,\mathcal{S}}$ maps the $\Zh$ states
$\{\mathit{init}, \mathit{awaiting}, \mathit{done}\}$ to the
$\Zw$ states $\{\mathit{closed}, \mathit{open}, \mathit{closed}\}$.
\end{itemize}

\paragraph{Tightness.}
The leakage is the additional information a $\Zw$-observer learns
about the $\Zh$-session beyond what is visible to a $\Zb$-observer
of the underlying $\Zb$ session. The two additional wire bytes are
\texttt{OPEN} and \texttt{CLOSE} markers, both of which are known
constants determined by the $\Zh$-message structure (any
$\Zh$-observer at the $\Zb$ layer would observe identical framing
because $\Zh$ already serializes through $\Zw$-style framing at
$\Zb$). The distribution of the $\Zw$-trace given a $\Zh$-trace is
deterministic, hence statistical distance zero, hence $\rho = 0
\le \mathrm{negl}(\lambda)$.

\paragraph{Property preservation.}
The security predicates of $\Zh$ are: confidentiality, authenticity,
post-quantum key derivation, channel-binding (when paired with
$\Zr$). All four predicates are valuated identically on the $\Zw$
state because the $\Zw$ session keys are derived from the same $\Zb$
key schedule. Hence
$\Pi(\varphi_{hw}) = \{\mathrm{conf}, \mathrm{auth}, \mathrm{pq},
\mathrm{cb}\}$.
\end{proof}

\subsection{$\varphi_{fw} : \Zf \hookrightarrow \Zw$}

\begin{proposition}[Proposition~5.2 of companion]
\label{prop:fw-proof}
The embedding $\varphi_{fw}$ that maps a FIX session to a single
$\Zw$ stream with FIX sequence numbers preserved as $\Zw$ envelope
sequence numbers is tight.
\end{proposition}

\begin{proof}
The map sends each FIX message $m$ with sequence number $n$ to a
$\Zw$ envelope $\langle m, n \rangle$. The $\Zw$ session is opened
at FIX-Logon and closed at FIX-Logout. Sequence-number gap-fill is
realized by $\Zw$ \texttt{REPLAY}/\texttt{ACK} framing.

\paragraph{Tightness.}
The wire-level activity of $\varphi_{fw}$ is exactly the activity a
$\Zf$-observer would see at the $\Zb$ layer, because $\Zf$ already
specifies its messages as $\Zb$ envelopes with FIX semantics. The
$\Zw$ envelope sequence number is the FIX sequence number,
verbatim. Statistical distance is zero, $\rho = 0$.

\paragraph{Property preservation.}
$\Zf$ inherits all $\Zb$ security predicates: confidentiality,
authenticity, post-quantum, channel-binding. The FIX-specific
predicate ``in-order delivery up to gap-fill'' is preserved by
$\Zw$ because $\Zw$ guarantees in-order delivery within a stream.
Hence $\Pi(\varphi_{fw}) \supseteq \{\mathrm{conf}, \mathrm{auth},
\mathrm{pq}, \mathrm{cb}, \mathrm{in\text{-}order}\}$.
\end{proof}

\subsection{$\varphi_{rb} : \Zr \hookrightarrow \Zb$}

\begin{proposition}[Proposition~5.3 of companion]
\label{prop:rb-proof}
The embedding $\varphi_{rb}$ that injects an RNS-bound identity into
the \texttt{identity} field of the $\Zb$ envelope at handshake time
is tight.
\end{proposition}

\begin{proof}
$\Zr$ is the simplest case: $\Zr$'s state is exactly $\Zb$'s state
with the \texttt{identity} field bound to the RNS-resolved
identifier, and $\Zr$'s message alphabet is a singleton extension of
$\Zb$'s (the channel-binding handshake message). The embedding is
$\varphi_{rb,\Sigma}(m) = (m')$ where $m'$ is the $\Zb$ envelope
with the \texttt{identity} field set, and
$\varphi_{rb,\mathcal{S}}(s) = s$ on the underlying state.

\paragraph{Tightness.}
The only wire-level difference is the population of the
\texttt{identity} field. From a $\Zb$-observer's perspective, the
\texttt{identity} field is part of the envelope; populating it does
not change the framing or the byte-count distribution. The
distribution of the $\Zb$-trace given a $\Zr$-trace is determined by
the RNS resolution (which is deterministic given the public identity
input), hence $\rho = 0$.

\paragraph{Property preservation.}
$\Zr$'s contribution is the channel-binding predicate $\mathrm{cb}$.
This is preserved because the binding token is in the $\Zb$
envelope; any future protocol layered on top of this $\Zb$ session
inherits the binding. Confidentiality, authenticity, and
post-quantum predicates are unchanged.
\end{proof}

\subsection{$\varphi_{mw} : \Zm \hookrightarrow \Zw$}

\begin{proposition}[Proposition~5.4 of companion]
\label{prop:mw-proof}
The embedding $\varphi_{mw}$ that maps an MCP tool-call/response
pair to a $\Zw$ stream where the request is a single envelope and
the response is a finite sequence of envelopes terminated by a
final-frame marker is tight.
\end{proposition}

\begin{proof}
The map sends each tool-call envelope $\mathit{call}$ to a singleton
$\Zw$ trace and each streaming response chunk $\mathit{chunk}_i$ to
a $\Zw$ envelope, terminated by a $\mathtt{FIN}$ marker.

\paragraph{Tightness.}
Streaming responses are natively a $\Zw$ feature; the final-frame
marker is a $\Zw$-level construct. The $\Zw$-trace given an
$\Zm$-trace is deterministic and identical in length and byte
content to what a $\Zb$-observer would see. $\rho = 0$.

\paragraph{Property preservation.}
$\Zm$'s tool-call authorization predicate is layered on top of $\Zb$
authentication. The four $\Zb$ predicates are inherited via $\Zw$
(by composition with \cref{prop:hw-proof}'s analysis specialized to
streaming). Hence
$\Pi(\varphi_{mw}) \supseteq \{\mathrm{conf}, \mathrm{auth},
\mathrm{pq}, \mathrm{cb}\}$.
\end{proof}

\subsection{$\varphi_{am} : \Za \hookrightarrow \Zm$}

\begin{proposition}[Proposition~5.5 of companion]
\label{prop:am-proof}
The embedding $\varphi_{am}$ that maps an A2A task to an MCP tool
call with structured arguments and an MCP streaming response
carrying task progress and result is tight.
\end{proposition}

\begin{proof}
A2A tasks are specified by a triple $(\mathit{task\_id}, \mathit{args},
\mathit{caps})$. The embedding maps this to an MCP tool call
$\mathit{call}(\mathit{tool} = \mathtt{a2a.execute},
\mathit{args} = (\mathit{task\_id}, \mathit{args}, \mathit{caps}))$
with structured arguments. Task progress events become MCP streaming
chunks; the final result is the terminal chunk before
$\mathtt{FIN}$.

\paragraph{Tightness.}
The mapping is structural: each A2A field is carried in a named MCP
argument. There is no wire-level activity in the MCP layer that is
not specified by the A2A trace. $\rho = 0$.

\paragraph{Property preservation.}
A2A inherits MCP's tool-call authorization (every A2A task is an
authorized MCP tool call) and, transitively, $\Zb$'s confidentiality,
authenticity, post-quantum, and channel-binding predicates.
\end{proof}

\subsection{Composite tightness}

\begin{corollary}[Composite tightness]
\label{cor:composite-tight}
The composite embeddings $\varphi_{mw} \circ \varphi_{am} : \Za
\hookrightarrow \Zw$ and $\varphi_{hw} \circ \varphi_{??} \circ
\cdots : \Za \hookrightarrow \Zb$ are tight.
\end{corollary}

\begin{proof}
By \cref{thm:cat-axioms}, leakage adds: $\rho_{\psi \circ \varphi}
= \rho_\psi + \rho_\varphi$. Since each constituent embedding has
$\rho = 0$, the composite has $\rho = 0$. Tightness is preserved.
\end{proof}

\section{Worked Example: PQ Inheritance for FIX-over-RNS}
\label{sec:fix-rns-example}

\begin{example}[Inherited PQ properties]
\label{ex:fix-rns-pq}
We make precise the claim of the companion paper that running $\Zf$
over $\Zr$ over $\Zw$ over $\Zb$ inherits the post-quantum
hybrid-KEM key schedule of $\Zb$ and the identity-binding guarantee
of $\Zr$.

By \cref{prop:fw-proof}, $\Zf \hookrightarrow \Zw$ is tight and
preserves $\mathrm{pq}$ (the post-quantum predicate). By
\cref{prop:hw-proof}'s argument applied to $\Zw \hookrightarrow
\Zb$ (the natural inclusion), $\Zw$ inherits $\mathrm{pq}$ from
$\Zb$. By \cref{prop:rb-proof}, $\Zr$ contributes $\mathrm{cb}$
(channel-binding) at the $\Zb$ envelope.

Composing,
\[
\Zf \xhookrightarrow{\varphi_{fw}, \mathrm{tight}} \Zw
\xhookrightarrow{\mathrm{tight}} \Zb,
\]
and the $\Zr$ binding sits at the $\Zb$ envelope. By
\cref{cor:stack-proof}, the bound for $\Zf$ inheriting both
$\mathrm{pq}$ and $\mathrm{cb}$ is
\[
\epsilon_{\Zf}^{\mathrm{pq}} \le \epsilon_{\Zb}^{\mathrm{pq}} +
\rho_{\varphi_{fw}} + \delta_{\varphi_{fw}} +
\rho_{\mathrm{Zw \hookrightarrow Zb}} + \delta_{\mathrm{Zw
\hookrightarrow Zb}} = \epsilon_{\Zb}^{\mathrm{pq}} +
\mathrm{negl}(\lambda),
\]
and similarly for $\mathrm{cb}$. Concretely, if $\Zb$'s hybrid KEM
gives $\epsilon_{\Zb}^{\mathrm{pq}} = 2^{-128}$ against the
modeled adversary class, then $\Zf$ over the stack achieves
$\epsilon_{\Zf}^{\mathrm{pq}} \le 2^{-128} + \mathrm{negl}(\lambda)$,
which for any reasonable $\lambda$ is well within the security
target.
\end{example}

\section{Discussion of Limitations}
\label{sec:limits}

The framework assumes:

\begin{enumerate}
\item \emph{LTS abstraction faithfulness.} Real implementations may
violate the LTS model (timing side-channels, cache effects). The
framework gives no guarantees against attacks below the LTS level.
\item \emph{Class-bounded adversaries.} The composition theorem is
parameterized over an adversary class $\mathcal{C}$. Quantum
adversaries against classical predicates require an explicit lift;
we have set up the framework so that the lift is mechanical (replace
$\mathcal{C}$ with a quantum-equipped class throughout) but the
underlying $\Zb$ proofs must hold against the lifted class for the
result to apply.
\item \emph{Tight embedding by construction.} The five concrete
embeddings are tight by inspection. A future sub-protocol whose
embedding is not tight would weaken the bound by the embedding's
leakage, which the framework computes explicitly. We do not commit
to all future sub-protocols being tight.
\item \emph{Availability is not modeled.} A denial-of-service attack
on $\Zb$ denies service to every protocol in the family. This is a
characteristic of the layered design, not a weakness of the
framework.
\end{enumerate}

\section{Conclusion}

We have proven the three foundational theorems of the categorical
framework: the category and 2-category axioms hold for $\CZ$, $\Zb$
is initial, and security properties compose along tight embeddings
with explicit and tight bounds. The five concrete embeddings used
by the family are tight. The composite stack
$\Za \hookrightarrow \Zm \hookrightarrow \Zw \hookrightarrow \Zb$
inherits the security properties of $\Zb$ up to a negligible
additive term. Future work includes machine-checked formalization
in Coq and extension to a quantum-equipped adversary class.

\printbibliography

\end{document}
